---
title: "INDIVIDUAL Vs. CORPORATION Loan Amount"
output: html_document
date: "2023-04-18"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = TRUE, fig.height = 3)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(broman)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
theme_set(theme_minimal())


```



## Introduction
Small businesses continue to play a critical role in driving economic growth and job creation in the United States, understanding the factors that determine the amount of funding they receive can be crucial for policymakers, lenders, and entrepreneurs alike. By leveraging loan-level data from the US Small Business Administration's 504 lending program, we aim to shed light on how the type of business and the year in which it received a loan can impact the amount of financing obtained. Our primary objective in analyzing this loan-level data is to gain insights into which types of small businesses are receiving the highest levels of funding and whether or not these funding levels have changed over time.The aim of this report is to investigate the extent to which the loan amount varies between different business types and across different years, by analyzing loan-level data from the US Small Business Administration's 504 lending programs. Through this analysis, we hope to identify patterns and trends that can inform lending practices and policies

## Background
#### (About the raw data)
The United States Small Business Administration’s most common loan program provides financial help for businesses that meet certain size and profitability requirements. These loans are long-term, fixed-rate financing of up to $5 million for major fixed assets. In each row the dataset is a different loan given to a small business in the United States via a lending program started by the United States Small Business Administration. The entries, updated quarterly, indicate each loan’s amount, term, interest rate, approval date, and current status; borrower type, name, and location; bank name and location; and more.  

#### (Data we are using)
Specifically, we will be focusing on loans issued between the years 2000 and 2002 and analyzing how the loan amount varies across different business types and over time. By narrowing our focus to this specific time period, we can more accurately identify patterns and trends in lending practices and policies that have emerged in during these years. The loan amount is a particularly important variable to consider, as it can significantly impact a small business's ability to grow and succeed. By analyzing this data, we hope to provide insights into how small businesses can better access the funding they need to thrive, and inform policymakers and lenders about potential gaps or disparities in lending practices.

#### Data Sitation
7(a) & 504 FOIA - FOIA - 504 (FY1991-FY2009) asof 221231.csv - U.S. Small Business Administration (SBA) | Open Data. (2023) https://data.sba.gov/dataset/7-a-504-foia/resource/ed462be9-e44f-45c6-818d-fca9b68cfeeb 

```{r  echo = FALSE}
baseData = read_csv("Project.csv")
baseData = baseData%>%
  filter(ApprovalFiscalYear >= 2000 & ApprovalFiscalYear <= 2002)%>%
  select(ApprovalFiscalYear, GrossApproval, BusinessType)%>%
  drop_na%>%
  group_by(BusinessType)%>%
  filter(BusinessType == "INDIVIDUAL" | BusinessType == "CORPORATION")


  
```

### Analysis
> In this dataset, we have information on two different types of businesses: Corporations and Individuals. We will analyze the total number of businesses for each type and the total amount of money they received overall.

```{r echo = FALSE}
Num_buiss= baseData %>%
  group_by(BusinessType)%>%
  summarize(Number_of_Buissnesses = n(),
            Total = sum(GrossApproval))
Num_buiss



plot1 = ggplot(Num_buiss, aes(x= BusinessType, y =Total )) +
  geom_col()
plot1

```
>This graph shows us that Corporations recieved a much greater total amount of money when compared to individual buisnesses, however, there is also a much greater amount of Corporations than Individual Buisinesses. To obtain these numbers we added up the total number of buisness for each type along with adding up the total loan amount they were approved for. We will now look at the proportional loan amount that each one of these buisnesses recieved, to determine if this great difference is primarly becuase of the difference in number of buissness rather than how much they are being aprroved for.


Is the total amount of money received proportional to the number of coorperations and number of individual busisness


Find the means for each BuisnessType and the difference
```{r  echo = FALSE}
mean_Approval = baseData%>%
  group_by(BusinessType)%>%
  summarize(n = n(),
            xbar = mean(GrossApproval),
            s = sd(GrossApproval))
mean_Approval


```

>There seems to be a very distint difference in the average amount of money that corporations recieved and the indivual buisnesses recieved. We are able to see that corporation on average recieved $448,756.20 while individual business recieved 294,396.10 dollars. We are also able to see that the standard deviation for amount receieved from corporations is also much larger than the standard deviation that the individual buisness received.  We can not rule out that this did not happen by chance though.  We will conduct a Hypothsis t-test difference in menas inorder to look at if this difference in average total loan amount difference or greater could happen by chance alone. 

> Is there a difference in the mean loan amount between cooperation or Individuals? (Hypothsis test t- test difference in means)

>Hypthesis:
Ho: Mu = 0
Ha: Mu != 0



Graph of the mean sampling distribution
```{r echo = FALSE}
ggplot(baseData, aes(x = GrossApproval, fill = BusinessType)) +
  geom_density(alpha = 0.5) +
  labs(title = "Sampling Distribution of Gross Approval Amount",
       x = "Gross Approval",
       y = "Density") +
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("INDIVIDUAL", "CORPORATION")) +
  theme_minimal()
```



```{r echo = FALSE}
corporation = baseData %>% 
  filter(BusinessType == "CORPORATION") 
individual = baseData %>% 
  filter(BusinessType == "INDIVIDUAL")
  
t.test(corporation$GrossApproval,individual$GrossApproval)
```
This a Welch Two Sample t-test on the Gross Approval amounts for Corporations and Individuals. The purpose of this test is to compare the means of two independent groups to determine if there is a significant difference between them. The t-test result shows a t-value of 25.135 and a degrees of freedom (df) of 2946. The p-value is extremely small (< 2.2e-16), indicating that there is a statistically significant difference between the means of the Gross Approval amounts for Corporations and Individuals.

The alternative hypothesis states that the true difference in means is not equal to 0, and given the results, we have evidence to support this claim. The 95% confidence interval for the difference in means ranges from 142,318.7 to 166,401.5. 


    - $H_0: \mu_C - \mu_I = 0$
    - $H_a: \mu_C - \mu_I \neq 0$


The sample of gross approval per business type follows a right skewed distribution with unknown population mean and standard deviation under the null hypothesis
    
We will use the t-statistic to test the hypotheses.

Under the null hypothesis, the t-statistic follows a t-distribution with 19 degrees of freedom!!!

Interpretation: The p-value is extremely small which indicates that we have strong evidence to reject the null hypothesis and suggest the alternative hypothesis as the true difference in means of gross loan approval amount between the two business types.

We will now construct a confidence interval to further explore this conclusion.

```{r echo = FALSE}
n1 = as.numeric(mean_Approval[1,2])
n2 = as.numeric(mean_Approval[2,2])
xbar1 = as.numeric(mean_Approval[1,3])
xbar2 = as.numeric(mean_Approval[2,3])
s1 = as.numeric(mean_Approval[1,4])
s2 = as.numeric(mean_Approval[2,4])
n1
se = sqrt((s1^2/n1) + (s2^2/n2))
df = n1 + n2 - 2
df
t_stat = (xbar1 - xbar2) / se
tcrit = qt(0.975, df)
lower = (xbar1 - xbar2) - tcrit*se
upper = (xbar1 - xbar2) + tcrit*se
cat("The 95% confidence interval for the difference in means is  (",round(lower, 2), ",",round(upper, 2), ")")
```
This is a 95% confidence interval for the difference in means of the gross loan approval amount between the two business types (Corporations and Individuals). It assumes that the sample of gross approval per business type follows a right-skewed distribution with unknown population mean and standard deviation under the null hypothesis.

The t-statistic is used to test the hypotheses, and under the null hypothesis, it follows a t-distribution with 19 degrees of freedom. Since the p-value is extremely small, there is strong evidence to reject the null hypothesis and support the alternative hypothesis, which claims that there is a true difference in the means of gross loan approval amounts between the two business types.

We are 95% confident that the mean difference in gross loan amount approved between the two business types is between \$142,463.9 and $166,528.6


## Discsion
In our analysis of the US Small Business Administration's 504 lending program data between 2000 and 2002, we investigated the extent to which the loan amount varies between different business types (Corporations and Individuals) and across different years. Our primary objective was to gain insights into which types of small businesses received the highest levels of funding and whether or not these funding levels changed over time.

Our analysis revealed that Corporations received a much greater total amount of money compared to Individual businesses during the studied period. However, there were also a greater number of Corporations than Individual businesses. To account for this discrepancy, we focused on the proportional loan amount that each business type received. We found that there is a significant difference in the average loan amount approved for Corporations and Individuals. Corporations received, on average, $448,756.20, while Individual businesses received $294,396.10.

To further support our findings, we conducted a Welch Two Sample t-test on the Gross Approval amounts for Corporations and Individuals. The test results showed a statistically significant difference between the means of the Gross Approval amounts for the two business types. We also constructed a 95% confidence interval for the difference in means, which provided an estimated range of $142,463.9 to $166,528.6 for the true difference in means of gross loan approval amounts between Corporations and Individuals.

### Limitations and Future Directions
While our analysis provides valuable insights, there are some potential shortcomings:

The data we used is limited to the years 2000 to 2002. A more comprehensive analysis could include data from more recent years or a broader range of years to understand the changes in lending practices over time.

We focused only on the 504 lending program, but there are other loan programs available for small businesses. Future analyses could include other lending programs to gain a more comprehensive understanding of funding patterns for small businesses.

The analysis does not account for other factors that might influence loan approval amounts, such as the industry of the business, the credit history of the borrower, or regional differences in lending practices. Future research could incorporate these factors to provide a more nuanced understanding of the differences in loan amounts between business types.

Our analysis could be further refined by examining the distribution of loan amounts within each business categroies. This could help identify potential disparities in funding levels for different types of businesses within each category.

### Inconclusion
In conclusion, our analysis of the US Small Business Administration's 504 lending program data for the years 2000 to 2002 revealed a significant difference in the mean loan amount approved for Corporations and Individuals. This finding underscores the importance of understanding the factors that determine the amount of funding received by different types of small businesses, as it can significantly impact their ability to grow and succeed.

The observed differences in loan amounts between Corporations and Individuals may have various implications for policymakers, lenders, and entrepreneurs alike. It is crucial to recognize the potential disparities in access to funding for different types of businesses and to consider measures to address these disparities to ensure a level playing field for all business types. 


In summary, our primary conclusions are that there is a significant difference in the mean loan amount approved for Corporations and Individuals, and that this difference could have a meaningful impact on the growth and success of these businesses. By addressing the disparities in access to funding and implementing policies that ensure equitable access for all small businesses, we can foster a thriving and inclusive entrepreneurial ecosystem that drives economic growth and job creation in the United States.
